Accelerated evolution of convective simulations

### 2018-07-30 referee_process/second_submission/referee_response.tex

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[aps, 11pt, singlecolumn]{revtex4-1} % Set the font size (10pt, 11pt and 12pt) and paper size (letterpaper, a4paper, etc)
\usepackage{natbib}
\bibliographystyle{apsrev}
\usepackage{setspace}

\begin{document}
\newenvironment{myquotation}{
\begin{quotation}
\itshape
}{ 
\end{quotation}
}
%----------------------------------------------------------------------------------------
%	LETTER CONTENT
%----------------------------------------------------------------------------------------
\noindent
Dear Eckart, other PRFluids editors, and anonymous referees,
$\,$\newline

\begin{singlespace}
We return our revised manuscript, ``Accelerated evolution in convective simulations,''
clarifying the paper where requested by the report of referee 1. 
Aside from these changes, the 
final pargaraph of the introductory section of the 
manuscript now contains references to the works of
Julien et al 1998 (ref 16) and Sprague et al 2006 (ref 17), whose techniques in
rapidly rotating, asymptotically reduced convection 
we felt should be mentioned in this work. A few grammatical and mechanical
corrections have been made throughout the manuscript.

We appreciate the very careful work of the referees in improving the clarity of
this manuscript. Below we include our responses to the report; the original
text of the report is included in italics, and our response follows inline in
unitalicized blocks.

### 2018-07-30 tex/bvp_paper.tex
editorial stuff

\begin{abstract}
High Peclet number, turbulent convection is a classic system with a large timescale
separation between flow speeds and the thermal relaxation time.
In this paper, we present a method of achieving Accelerated Evolution (AE) of convective simulations which
fast-forwards through the long thermal evolution of these systems, and we test the
validity of this method. The AE procedure involves measuring the dynamics of convection
early in a simulation and using its characteristics to adjust the mean thermodynamic
profile within the domain towards its evolved state. We study \RB convection as a test case for AE.  
Evolved flow properties of AE solutions are measured to be within a few percent
of solutions which are reached through Standard Evolution (SE) over a full thermal timescale.
At the highest values of the Rayleigh number at which we compare SE and AE,
we find that AE solutions require roughly an order of magnitude fewer cpu-hours
to compute than SE solutions.
\end{abstract}
\maketitle

@@

The vast difference between convective and thermal timescales has long plagued
numericists studying convection, and an abundance of approaches has been employed to
study thermally relaxed solutions. One popular method for accelerating the convergence
of high-Ra solutions is by ``bootstrapping'' -- the process of using the flow
fields in a converged solution at low Ra as initial conditions for a simulation at high
Ra.  This method has been used with great success \cite{johnston&doering2009, verzicco&camussi1997},
but it is not without its faults.  In systems in which there are multiple
stable solutions, such as a roll state and a shearing state of convection,
the large-scale convective structures present in the
low Ra solution imprint onto the dynamics of the new, high Ra solution. 
It is possible that this puts the high Ra solution into a different stable state
than it would naturally reach from an initial, hydrostatically stable configuration.
Another commonly-used tactic in
moderate-Ra simulations is to first solve a simple model of the system in question,
and then use the solution of that model as initial conditions for full nonlinear
direct numerical simulations. For example, past studies have used initial conditions from
a linear eigenvalue solve \cite{hurlburt&all1984} in plane-parallel studies, or
axisymmetric solutions in studies of convection in 3D cylinders \cite{verzicco&camussi1997}. 
In other systems, particularly when convective zones are adjacent to stable regions,
authors often choose initial conditions which are not in a classic, hydrostatic,
conductive state. Rather, either through knowledge of low-Ra solutions \cite{couston&all2017}
or broader convective theories such as Mixing Length Theory
\cite{brandenburg&all2005}, initial conditions can be adjusted such that the stratification
within the convective domain is closer to a relaxed state than the largely
unstable hydrostatic state.

@@

In this work, we study a method of achieving accelerated evolution of
convective simulations. 
Our technique is similar in essence to the approach used in asymptotically 
reduced models of rapidly rotating convection, in which the
slowly evolving mean temperature profile is timestepped separately from the
fast convective dynamics \cite{julien&all1998, sprague&all2006}. 
We couple measurements of the dynamics of unstable, evolving
convective simulations with knowledge about energy balances in the desired solution
to instantaneously and self-consistently adjust the mean vertical thermodynamic 
profile toward its relaxed state. 
While a technique of this kind has been used by many studies previously 
(e.g., \cite{hurlburt&all1986}),
the details of implementation, the convergence properties, and whether or not the
thermally relaxed state achieved from 
accelerated evolution corresponds to the relaxed state in standard evolution 
are not documented.
In section \ref{sec:experiment}, we describe our convective simulations and
numerical methods. In section \ref{sec:ae}, we describe the procedure for 
achieving accelerated evolution. In
section \ref{sec:results}, we compare accelerated evolution solutions
to solutions obtained from standard evolution through a full thermal diffusion timescale. 
In section \ref{sec:speedups}, we compare the numerical cost of accelerated
solutions to standard solutions.
Finally,
in section \ref{sec:extensions}, we offer concluding remarks and
discuss extensions of the methods presented here.

@@

\section{Experiment}
\label{sec:experiment}
In this work we study a simple form of thermal convection:
incompressible \RB convection under the Oberbeck-Boussinesq approximation,
such that the fluid
has a constant kinematic viscosity ($\nu$), thermal diffusivity ($\kappa$), and coefficient
of thermal expansion ($\alpha$). The density of the fluid is a constant, $\rho_0$,
except in the buoyancy term, where it is $\rho = \rho_0(1  - \alpha T_1)$.
The gravitational acceleration, $\bm{g} = - g\bm{\hat{z}}$, is constant.
The equations of motion are \cite{spiegel&veronis1960}:
\begin{gather}
\DivU = 0, 
	\label{eqn:incompressible}
\\
\frac{\partial \bm{u}}{\partial t} + \bm{u}\cdot\grad\bm{u} =
-\frac{1}{\rho_0}\grad P - g( 1 - \alpha T_1)\hat{z} + \nu\grad^2\bm{u}, 
	\label{eqn:dim_bouss_momentum}
\\
\frac{\partial T_1}{\partial t} + \bm{u}\cdot\grad(T_0 + T_1) = \kappa\grad^2 T_1,
	\label{eqn:dim_bouss_energy}
\end{gather}
where $\bm{u} = u\bm{\hat{x}} + v\bm{\hat{y}} + w\bm{\hat{z}}$ is the velocity, 
$T = T_0(z) + T_1(x, y, z, t)$ are the initial and fluctuating components of temperature, 
and $P$ is the kinematic pressure. The initial temperature profile, $T_0$, decreases
linearly with height.
We non-dimensionalize these equations such that
length is in units of the layer height ($L_z$),
temperature is in units of the initial temperature jump across the layer ($\Delta T_0 = L_z \partial_z T_0$), 
and velocity is in units of the freefall velocity ($v_{\text{ff}} = \sqrt{\alpha g L_z^2 \partial_z T_0}$).
By these choices, one time unit is a freefall time ($t_{\text{ff}} = L_z/v_{\text{ff}}$).
We introduce a reduced kinematic pressure,
$\varpi \equiv (P / \rho_0 + \phi + |\bm{u}|^2 / 2) / v_{\text{ff}}^2$, where the gravitational
potential, $\phi$, is defined such that $\bm{g} = -\grad \phi$. 
In non-dimensional form, and substituting 
$\bm{u}\cdot\grad\bm{u} = \grad(|\bm{u}|^2/2) - \bm{u}\times(\grad\times\bm{u})$
and $\grad^2\bm{u} = -\grad\times(\grad\times\bm{u})$, Eqns.~\ref{eqn:dim_bouss_momentum} \& \ref{eqn:dim_bouss_energy}
become
\begin{align}
\frac{\partial \bm{u}}{\partial t} + \grad \varpi - T_1\hat{z} + \mathcal{R}\grad\times\bm{\omega} &= \bm{u}\times\bm{\omega},
	\label{eqn:bouss_momentum}
\\
\frac{\partial T_1}{\partial t} - \mathcal{P}\grad^2 T_1 + w \frac{\partial T_0}{\partial z} &= - \bm{u}\cdot\grad T_1,
	\label{eqn:bouss_energy}
\end{align}
where $\bm{\omega} = \grad \times \bm{u}$ is the vorticity.
The dimensionless control parameters $\mathcal{R}$ and $\mathcal{P}$ 
are set by the Rayleigh (Ra) and Prandtl (Pr) numbers,
\begin{equation}
\mathcal{R} \equiv \sqrt{\frac{\text{Pr}}{\text{Ra}}}, \qquad \mathcal{P} \equiv \frac{1}{\sqrt{\text{Pr}\,\text{Ra}}}, \qquad
\text{Ra} = \frac{g \alpha L_z^4 \partial_z T_0}{\nu\kappa} = \frac{(L_z\,v_{\text{ff}})^2}{\nu\kappa}, \qquad \text{Pr} = \frac{\nu}{\kappa}.
\end{equation}
We hold Pr $= 1$ constant throughout this work, such that $\mathcal{P} = \mathcal{R}$.
$\mathcal{P}$ and $\mathcal{R}$ are related to the inverse Reynolds and Peclet numbers of the
evolved flows.

In Eqns.~(\ref{eqn:incompressible}), (\ref{eqn:bouss_momentum}), \& (\ref{eqn:bouss_energy}),
linear terms are grouped on the left-hand side of the equations, while nonlinear terms
are found on the right-hand side. We timestep linear terms implicitly, and nonlinear
terms explicitly.
We utilize the 
Dedalus\footnote{\url{http://dedalus-project.org/}} 
pseudospectral framework \cite{burns&all2016} to evolve  
Eqns.~(\ref{eqn:incompressible}), (\ref{eqn:bouss_momentum}), \& (\ref{eqn:bouss_energy}) 
forward in time
using an implicit-explicit (IMEX), third-order, four-stage 
Runge-Kutta timestepping scheme RK443 \cite{ascher&all1997}. The code used to run the simulations
in this work is included in the supplemental materials as a zip file.

@@

\section{The method of Accelerated Evolution}
\label{sec:ae}
Here we describe a method of Accelerated Evolution (AE), which we use 
to rapidly evolve the thermodynamic state of convective simulations.  
We compare this AE method to Standard Evolution
(SE), in which we evolve the atmosphere from noise initial conditions
for one thermal diffusion time,
$t_\kappa = \mathcal{P}^{-1}$. Both AE and SE simulations begin with identical
initial conditions, as described in section \ref{sec:experiment}.
As Ra increases, and $\mathcal{P}$ decreases, SE solutions become intractable, 
while the timeframe of convergence for an AE solution remains nearly constant
in simulation freefall time units (see table \ref{table:run_parameters} in
appendix \ref{appendix:run_table}).

We study in depth a 2D simulation at $S = 10^5$ to demonstrate the power of AE.
We compare kinetic energy (KE, black line) and mean temperature (blue line)
traces from a SE run in Fig.~\ref{fig:time_trace}a to an AE run
in Fig.~\ref{fig:time_trace}c.
In Fig.~\ref{fig:time_trace}a, the time evolution of the SE simulation is shown.
The KE grows exponentially from white noise during the
first $\sim 25$ $t_{\text{ff}}$. The solution then saturates and begins to slowly
relax toward the saturated isothermal profile in the interior of the domain.
This slow relaxation is evident in the behavior of the blue line, which measures
$\angles{T} - T_{\text{top}}$, where \angles{T} is the volume-average of $T$, and
$T_{\text{top}} = -0.5$ is the temperature at the upper boundary.
The mean atmospheric temperature and
kinetic energies are fully converged when $t = 4000$ $t_{\text{ff}}$ $= 0.35$ $t_{\kappa}$.
We show roughly the first thousand freefall time
units of evolution, as well as the evolved thermodynamic state reached after a full
thermal time of evolution.  In Fig.~\ref{fig:time_trace}c, similar traces are
shown for the corresponding AE solution
at the same parameters. The same linear growth phase occurs, but shortly after
the peak of convective transient we accelerate the convergence of the atmosphere
through the process which we describe below. We adjust the 1D vertical profile of
the atmosphere three times, as denoted by the three labeled
arrows in the graph numbered 1-3.  The third profile adjustment associated with
arrow 3 is small enough (see appendix \ref{appendix:recipe}) that we assume
the atmosphere is sufficiently converged, and we begin to sample the evolved
convective dynamics.

The horizontally averaged profiles of the vertical conductive flux, 
F$_{\kappa} = \angles{-\kappa\grad(T_0 + T_1)}_{x,y}$, and the vertical convective enthalpy flux,
F$_{\text{E}} = \angles{w(T_0 + T_1)}_{x,y}$, are the basis of the AE method.
Here we use $\angles{}_{x,y}$ to represent a horizontal average. We measure
both of these fluxes early in a simulation, retrieving profiles such as
those shown in Fig.~\ref{fig:time_trace}b.
As the atmosphere evolves towards
the isothermal profile specified by the upper (cold) boundary condition, excess
temperature throughout the atmosphere must leave the domain. This excess thermal
energy leaves through the upper boundary, as seen in 
Fig.~\ref{fig:time_trace}b, where the amount of flux exiting at the top of the domain
is nearly 20 times larger the flux entering the bottom of the domain. Once the atmospheric
temperature profile reaches its evolved state, the flux entering the bottom boundary
is equal to the flux exiting through the upper boundary.  In general, this 
evolution is slow in SE (Fig.~\ref{fig:time_trace}a), but AE (Fig.~\ref{fig:time_trace}c)
can rapidly advance a system whose fluxes are in a strongly disequilibrium state (Fig.~\ref{fig:time_trace}b),
into a near-equilibrium state, as shown in Fig.~\ref{fig:time_trace}d. In
this final state both boundaries conduct the same amount of flux.
The converged fluxes achieved through AE are at most 5\% different from the SE solution, 
as shown in Fig.~\ref{fig:time_trace}e. 
This is a very small difference considering the short timescales on which convergence
is reached and the strongly disequilibrium state used to inform the AE process.

@@

To reduce the conductive flux by $\xi$, we examine the 
time-stationary state, horizontally- and time-averaged
Eqns.~(\ref{eqn:bouss_momentum}) and (\ref{eqn:bouss_energy}); after 
neglecting terms which vanish due to symmetry, these equations become
\vspace{-0.2cm}
\begin{gather}
\frac{\partial}{\partial z}\angles{\varpi}_{x,y} - \angles{T_1}_{x,y}\hat{z} = \angles{\bm{u}\times\bm{\omega}}_{x,y, \text{ ev}},
	\label{eqn:bouss_BVP_momentum}
\\
\frac{\partial}{\partial z}\text{F}_{\text{E, ev}} - \mathcal{P}\frac{\partial^2}{\partial z^2} \angles{T_1}_{x,y} = 0.
	\label{eqn:bouss_BVP_energy}
\end{gather}
Here, we construct  $\text{F}_{\text{E, ev}} = \xi \text{F}_{\text{E}}$ 
and $\angles{\bm{u}\times\bm{\omega}}_{x, y,\text{ ev}}
= \xi\angles{\bm{u}\times\bm{\omega}}_{x, y}$.
Using these profiles and 
Eqns. (\ref{eqn:bouss_BVP_momentum}) \& (\ref{eqn:bouss_BVP_energy}),
we solve for $\angles{\varpi}_{x,y}$ and $\angles{T_1}_{x,y}$.
Our choice of fixing the temperature at the top boundary and the flux at the bottom
boundary ensures that there is a unique solution for $\angles{T_1}_{x,y}$ given the evolved
fluxes. If flux were fixed at both boundaries, there would be infinitely many
temperature solutions. If temperature were fixed at both boundaries, 
$\text{F}_{\text{B}}$ would not be precisely known \emph{a priori}.

Solving the above equations adjusts the mean thermal state and 
$\text{F}_{\kappa}$ of the atmosphere while leaving $\text{F}_{\text{E}}$
unchanged. In the  
bulk of the domain, convective enthalpy flux dominates transport, and so there we
assume that $\text{F}_{\text{tot}} \approx \text{F}_{\text{E}}$. Upon 
removing terms which vanish due to symmetry, we note that the convective
enthalpy flux is carried solely by the velocity field and perturbations in
temperature away from the mean state, $\text{F}_{\text{E}} = 
\angles{w(T - \angles{T})}$.
In order to instantaneously decrease the convective enthalpy flux in a manner
which is consistent with the conductive flux, 
we multiply both the velocity, $\bm{u}$,
and temperature perturbations about the mean state, $T-\angles{T}_{x,y}$,
 by $\sqrt{\xi}$. This scaling of the velocity field is the reason that we
multiply $\angles{\bm{u}\times\bm{\omega}}_{x, y}$, which is nondimensionally
of order $u^2/L$, by $\xi$ while constructing
$\angles{\bm{u}\times\bm{\omega}}_{x, y,\text{ ev}}$.

@@

The Nusselt number (Nu) quantifies the efficiency of convective heat transport
and is defined as
\begin{equation}
\text{Nu} = \frac{\angles{F_{\text{conv}} + F_{\text{cond}}}}{\angles{F_{\text{cond, ref}}}}
 = \frac{\angles{wT - \mathcal{P}\partial_z T}}{\angles{- \mathcal{P} \partial_z T}},
\end{equation}
where the volume average of a quantity, $\eta$, is shown as $\angles{\eta}$.
The time evolution of Nu in AE and SE is compared to the mean temperature evolution
in Fig.~\ref{fig:nu_v_time}. In Fig.~\ref{fig:nu_v_time}a, we show the evolution
of the SE run at $S = 10^5$. The black trace shows the instantaneous value of
Nu, and the overplotted grey line shows a moving time average of Nu. The
time average is taken using a centered boxcar window whose width is 50 $t_{\text{ff}}$.
The mean value of Nu evolves towards its final value (dashed horizontal
grey line) rapidly compared to $\angles{T}-T_{\text{top}}$ (blue), 
which approaches its evolved state (dashed horizontal blue line) slowly.
As such, it is possible to measure the statistically stationary mean value of
Nu after only a few hundred simulated freefall times, 
as has been done previously \cite{stevens&all2010}. However, fluctuations in
Nu around the mean value at early times have both a larger magnitude and higher
frequency than the fluctuations in the relaxed state.
In Fig.~\ref{fig:nu_v_time}b, we show the AE traces of Nu and
$\angles{T}-T_{\text{top}}$ at $S = 10^5$. The times at which AE solves occur are marked
by arrows. Nu reaches its evolved mean value slightly more rapidly than in the SE case,
and the frequency and magnitude of fluctuations in Nu away from the mean resemble the
final relaxed state of SE.

@@

$10^{1/3}$	    &	$2.79 \cdot 10^3$       &	32$\times$64$\,\,\,\,\,\,\,$&	$52.8$	&	$340$	&	100	&	1.46	&	1.46	\\

>>>>

\begin{abstract}
High Peclet number, turbulent convection is a classic system with a large timescale
separation between flow speeds and the thermal relaxation time.
In this paper, we present a method of
fast-forwarding through the long thermal relaxation of convective simulations, and we test the
validity of this method. This Accelerated Evolution (AE) method involves measuring the dynamics of convection
early in a simulation and using its characteristics to adjust the mean thermodynamic
profile within the domain towards its evolved state. We study \RB convection as a test case for AE.  
Evolved flow properties of AE solutions are measured to be within a few percent
of solutions which are reached through Standard Evolution (SE) over a full thermal diffusion timescale.
At the highest values of the Rayleigh number at which we compare SE and AE,
we find that AE solutions require roughly an order of magnitude fewer cpu-hours
to compute than SE solutions.
\end{abstract}
\maketitle

@@

The vast difference between convective and thermal timescales has long plagued
numericists studying convection, and an abundance of approaches has been employed to
study thermally relaxed solutions. One popular method for accelerating the convergence
of high-Ra solutions is by ``bootstrapping'' -- the process of using the relaxed flow
fields and thermal structure of a low Ra solution as initial conditions for a simulation at high
Ra.  This method has been used with great success \cite{johnston&doering2009, verzicco&camussi1997},
but it is not without its faults.  In systems in which there are multiple
stable solutions, such as a roll state and a shearing state of convection,
the large-scale convective structures present in the
low Ra solution imprint onto the dynamics of the new, high Ra solution. 
It is possible that this puts the high Ra solution into a different stable state
than it would naturally reach from an initial, hydrostatically stable configuration.
Another commonly-used tactic in
moderate-Ra simulations is to first solve a simple model of the system in question,
and then use the solution of that model as initial conditions for full nonlinear
direct numerical simulations. For example, past studies have used initial conditions from
a linear eigenvalue solve \cite{hurlburt&all1984} in plane-parallel studies, or
axisymmetric solutions in studies of convection in 3D cylinders \cite{verzicco&camussi1997}. 
In other systems, particularly when convective zones are adjacent to stable regions,
authors often choose initial conditions which are not in a classic, hydrostatic,
conductive state. Rather, either through knowledge of low-Ra solutions \cite{couston&all2017}
or broader convective theories such as Mixing Length Theory
\cite{brandenburg&all2005}, initial conditions can be adjusted such that the stratification
within the convective domain is closer to a relaxed state than the largely
unstable hydrostatic state.

@@

In this work, we study a method of achieving accelerated evolution of
convective simulations. 
Our technique is similar in essence to the approach used in asymptotically 
reduced models of rapidly rotating convection, in which the
mean temperature profile evolves separately from the
fast convective dynamics \cite{julien&all1998, sprague&all2006}. 
We couple measurements of the dynamics of unstable, evolving
convective simulations with knowledge about energy balances in the relaxed solution
to instantaneously and self-consistently adjust the mean vertical thermodynamic 
profile toward its relaxed state. 
While a technique of this kind has been used by many studies previously 
(e.g., \cite{hurlburt&all1986}),
the details of implementation, the convergence properties, and whether or not the
thermally relaxed state achieved from 
accelerated evolution corresponds to the relaxed state in standard evolution 
are not documented.
In section \ref{sec:experiment}, we describe our convective simulations and
numerical methods. In section \ref{sec:ae}, we describe the procedure for 
achieving accelerated evolution. In
section \ref{sec:results}, we compare accelerated evolution solutions
to solutions obtained from standard evolution through a full thermal diffusion timescale. 
In section \ref{sec:speedups}, we compare the numerical cost of accelerated
solutions to standard solutions.
Finally,
in section \ref{sec:extensions}, we offer concluding remarks and
discuss extensions of the methods presented here.

@@

\section{Experiment}
\label{sec:experiment}
In this work we study a simple form of thermal convection:
incompressible \RB convection under the Oberbeck-Boussinesq approximation,
such that the fluid
has a constant kinematic viscosity ($\nu$), thermal diffusivity ($\kappa$), and coefficient
of thermal expansion ($\alpha$). The density of the fluid is a constant, $\rho_0$,
except in the buoyancy term, where it is $\rho = \rho_0(1  - \alpha T_1)$.
The gravitational acceleration, $\bm{g} = - g\bm{\hat{z}}$, is constant.
The equations of motion are \cite{spiegel&veronis1960}:
\begin{gather}
\DivU = 0, 
	\label{eqn:incompressible}
\\
\frac{\partial \bm{u}}{\partial t} + \bm{u}\cdot\grad\bm{u} =
-\frac{1}{\rho_0}\grad P - g( 1 - \alpha T_1)\hat{z} + \nu\grad^2\bm{u}, 
	\label{eqn:dim_bouss_momentum}
\\
\frac{\partial T_1}{\partial t} + \bm{u}\cdot\grad(T_0 + T_1) = \kappa\grad^2 (T_0 + T_1),
	\label{eqn:dim_bouss_energy}
\end{gather}
where $\bm{u} = u\bm{\hat{x}} + v\bm{\hat{y}} + w\bm{\hat{z}}$ is the velocity, 
$T = T_0(z) + T_1(x, y, z, t)$ are the initial and fluctuating components of temperature, 
and $P$ is the kinematic pressure. The initial temperature profile, $T_0$, decreases
linearly with height.
We non-dimensionalize these equations such that
length is in units of the layer height ($L_z$),
temperature is in units of the initial temperature jump across the layer ($\Delta T_0 = L_z \partial_z T_0$), 
and velocity is in units of the freefall velocity ($v_{\text{ff}} = \sqrt{\alpha g L_z^2 \partial_z T_0}$).
By these choices, one time unit is a freefall time ($t_{\text{ff}} = L_z/v_{\text{ff}}$).
We introduce a reduced kinematic pressure,
$\varpi \equiv (P / \rho_0 + \phi + |\bm{u}|^2 / 2) / v_{\text{ff}}^2$, where the gravitational
potential, $\phi$, is defined such that $\bm{g} = -\grad \phi$. 
In non-dimensional form, and substituting 
$\bm{u}\cdot\grad\bm{u} = \grad(|\bm{u}|^2/2) - \bm{u}\times(\grad\times\bm{u})$
and $\grad^2\bm{u} = -\grad\times(\grad\times\bm{u})$, Eqns.~\ref{eqn:dim_bouss_momentum} \& \ref{eqn:dim_bouss_energy}
become
\begin{align}
\frac{\partial \bm{u}}{\partial t} + \grad \varpi - T_1\hat{z} + \mathcal{R}\grad\times\bm{\omega} &= \bm{u}\times\bm{\omega},
	\label{eqn:bouss_momentum}
\\
\frac{\partial T_1}{\partial t} - \mathcal{P}\grad^2 T_1 + w \frac{\partial T_0}{\partial z} &= - \bm{u}\cdot\grad T_1,
	\label{eqn:bouss_energy}
\end{align}
where $\bm{\omega} = \grad \times \bm{u}$ is the vorticity.
The dimensionless control parameters $\mathcal{R}$ and $\mathcal{P}$ 
are set by the Rayleigh (Ra) and Prandtl (Pr) numbers,
\begin{equation}
\mathcal{R} \equiv \sqrt{\frac{\text{Pr}}{\text{Ra}}}, \qquad \mathcal{P} \equiv \frac{1}{\sqrt{\text{Pr}\,\text{Ra}}}, \qquad
\text{Ra} = \frac{g \alpha L_z^4 \partial_z T_0}{\nu\kappa} = \frac{(L_z\,v_{\text{ff}})^2}{\nu\kappa}, \qquad \text{Pr} = \frac{\nu}{\kappa}.
\end{equation}
We hold Pr $= 1$ constant throughout this work, such that $\mathcal{P} = \mathcal{R}$.
$\mathcal{P}$ and $\mathcal{R}$ are related to the inverse Reynolds and Peclet numbers of the
evolved flows.

In Eqns.~(\ref{eqn:incompressible}), (\ref{eqn:bouss_momentum}), \& (\ref{eqn:bouss_energy}),
linear terms are grouped on the left-hand side of the equations, while nonlinear terms
are found on the right-hand side. We timestep linear terms implicitly, and nonlinear
terms explicitly.
We utilize the 
Dedalus\footnote{\url{http://dedalus-project.org/}} 
pseudospectral framework \cite{burns&all2016} to evolve  
Eqns.~(\ref{eqn:incompressible}), (\ref{eqn:bouss_momentum}), \& (\ref{eqn:bouss_energy}) 
forward in time
using an implicit-explicit (IMEX), third-order, four-stage 
Runge-Kutta timestepping scheme RK443 \cite{ascher&all1997}. The code used to run the simulations
in this work is included in the supplemental materials as a zip file \cite{supp}.

@@

\section{The method of Accelerated Evolution}
\label{sec:ae}
Here we describe a method of Accelerated Evolution (AE), which we use 
to rapidly relax the thermodynamic state of convective simulations.  
We compare this AE method to Standard Evolution
(SE), in which we evolve the atmosphere from noise initial conditions
for one thermal diffusion time,
$t_\kappa = \mathcal{P}^{-1}$. Both AE and SE simulations begin with identical
initial conditions, as described in section \ref{sec:experiment}.
As Ra increases, and $\mathcal{P}$ decreases, SE solutions become intractable, 
while the timeframe of convergence for an AE solution remains nearly constant
in simulation freefall time units (see table \ref{table:run_parameters} in
appendix \ref{appendix:run_table}).

We study in depth a 2D simulation at $S = 10^5$ to demonstrate the power of AE.
We compare kinetic energy (KE, black line) and mean temperature (blue line)
traces from a SE run in Fig.~\ref{fig:time_trace}a to an AE run
in Fig.~\ref{fig:time_trace}c.
In Fig.~\ref{fig:time_trace}a, the time evolution of the SE simulation is shown.
The KE grows exponentially from white noise during the
first $\sim 25$ $t_{\text{ff}}$. The solution then saturates and begins to slowly
relax toward the saturated isothermal profile in the interior of the domain.
This slow relaxation is evident in the behavior of the blue line, which measures
$\angles{T} - T_{\text{top}}$, where \angles{T} is the volume-average of $T$, and
$T_{\text{top}} = -0.5$ is the temperature at the upper boundary.
The mean atmospheric temperature and
kinetic energies are fully converged when $t = 4000t_{\text{ff}} = 0.35t_{\kappa}$.
We show roughly the first thousand freefall time
units of evolution, as well as the evolved thermodynamic state reached after a full
thermal time of evolution.  In Fig.~\ref{fig:time_trace}c, similar traces are
shown for the corresponding AE solution
at the same parameters. The same linear growth phase occurs, but shortly after
the peak of convective transient we accelerate the convergence of the atmosphere
through the process which we describe below. We adjust the 1D vertical profile of
the atmosphere three times, as denoted by the three labeled
arrows in the graph numbered 1-3.  The third profile adjustment associated with
arrow 3 is small enough (see appendix \ref{appendix:recipe}) that we assume
the atmosphere is sufficiently converged, and we begin to sample the evolved
convective dynamics.

The horizontally averaged profiles of the vertical conductive flux, 
F$_{\kappa} = \angles{-\kappa\partial_z T}_{x,y}$, and the vertical convective enthalpy flux,
F$_{\text{E}} = \angles{wT}_{x,y}$, are the basis of the AE method.
Here we use $\angles{}_{x,y}$ to represent a horizontal average. We measure
both of these fluxes early in a simulation, retrieving profiles such as
those shown in Fig.~\ref{fig:time_trace}b.
As the atmosphere relaxes towards
the isothermal profile specified by the upper (cold) boundary condition, excess
temperature throughout the atmosphere must leave the domain. This excess thermal
energy leaves through the upper boundary, as seen in 
Fig.~\ref{fig:time_trace}b, where the amount of flux exiting at the top of the domain
is nearly 20 times larger the flux entering the bottom of the domain. Once the atmospheric
temperature profile reaches its evolved state, the flux entering the bottom boundary
is equal to the flux exiting through the upper boundary.  In general, this 
evolution is slow in SE (Fig.~\ref{fig:time_trace}a), but AE (Fig.~\ref{fig:time_trace}c)
can rapidly advance a system whose fluxes are in a strongly disequilibrium state (Fig.~\ref{fig:time_trace}b),
into a near-equilibrium state, as shown in Fig.~\ref{fig:time_trace}d. In
this final state both boundaries conduct the same amount of flux.
The converged fluxes achieved through AE are at most 5\% different from the SE solution, 
as shown in Fig.~\ref{fig:time_trace}e. 
This is a very small difference considering the short timescales on which convergence
is reached and the strongly disequilibrium state used to inform the AE process.

@@

To reduce the conductive flux by $\xi$, we examine the 
horizontally- and time-averaged
Eqns.~(\ref{eqn:bouss_momentum}) and (\ref{eqn:bouss_energy}) in the time-stationary state; after 
neglecting terms which vanish due to symmetry, these equations become
\vspace{-0.2cm}
\begin{gather}
\frac{\partial}{\partial z}\angles{\varpi}_{x,y} - \angles{T_1}_{x,y}\hat{z} = \angles{\bm{u}\times\bm{\omega}}_{x,y, \text{ ev}},
	\label{eqn:bouss_BVP_momentum}
\\
\frac{\partial}{\partial z}\text{F}_{\text{E, ev}} - \mathcal{P}\frac{\partial^2}{\partial z^2} \angles{T_1}_{x,y} = 0.
	\label{eqn:bouss_BVP_energy}
\end{gather}
Here, we construct  $\text{F}_{\text{E, ev}} = \xi \text{F}_{\text{E}}$ 
and $\angles{\bm{u}\times\bm{\omega}}_{x, y,\text{ ev}}
= \xi\angles{\bm{u}\times\bm{\omega}}_{x, y}$.
Using these profiles and 
Eqns. (\ref{eqn:bouss_BVP_momentum}) \& (\ref{eqn:bouss_BVP_energy}),
we solve for $\angles{\varpi}_{x,y}$ and $\angles{T_1}_{x,y}$.
Our choice of fixing the temperature at the top boundary and the flux at the bottom
boundary ensures that there is a unique solution for $\angles{T_1}_{x,y}$ given the evolved
fluxes. If flux were fixed at both boundaries, there would be infinitely many
temperature solutions. If temperature were fixed at both boundaries, 
$\text{F}_{\text{B}}$ would not be precisely known \emph{a priori}.

Solving the above equations adjusts the mean thermal state and 
$\text{F}_{\kappa}$ of the atmosphere while leaving $\text{F}_{\text{E}}$
unchanged. In the  
bulk of the domain, convective enthalpy flux dominates transport, and so there we
assume that $\text{F}_{\text{tot}} \approx \text{F}_{\text{E}}$. Upon 
removing terms which vanish due to symmetry, we note that the convective
enthalpy flux is carried solely by the velocity field and perturbations in
temperature away from the mean state, $\text{F}_{\text{E}} = 
\angles{w(T_1 - \angles{T_1})}$.
In order to instantaneously decrease the convective enthalpy flux in a manner
which is consistent with the conductive flux, 
we multiply both the velocity, $\bm{u}$,
and temperature perturbations about the mean state, $T_1-\angles{T_1}_{x,y}$,
 by $\sqrt{\xi}$. This scaling of the velocity field is the reason that we
multiply $\angles{\bm{u}\times\bm{\omega}}_{x, y}$, which is nondimensionally
of order $u^2/L$, by $\xi$ while constructing
$\angles{\bm{u}\times\bm{\omega}}_{x, y,\text{ ev}}$.

@@

The Nusselt number (Nu) quantifies the efficiency of convective heat transport
and is defined as
\begin{equation}
\text{Nu} = \frac{\angles{F_{\text{conv}} + F_{\text{cond}}}}{\angles{F_{\text{cond, ref}}}}
 = \frac{\angles{wT - \mathcal{P}\partial_z T}}{\angles{- \mathcal{P} \partial_z T}},
\end{equation}
where the volume average of a quantity, $\eta$, is shown as $\angles{\eta}$.
The time evolution of Nu in AE and SE is compared to the mean temperature evolution
in Fig.~\ref{fig:nu_v_time}. In Fig.~\ref{fig:nu_v_time}a, we show the evolution
of the SE run at $S = 10^5$. The black trace shows the instantaneous value of
Nu, and the overplotted grey line shows a moving time average of Nu. The
time average is taken using a centered boxcar window whose width is 50 $t_{\text{ff}}$.
The mean value of Nu evolves towards its relaxed value (dashed horizontal
grey line) rapidly compared to $\angles{T}-T_{\text{top}}$ (blue), 
which approaches its evolved state (dashed horizontal blue line) slowly.
As such, it is possible to measure the statistically stationary mean value of
Nu after only a few hundred simulated freefall times, 
as has been done previously \cite{stevens&all2010}. However, fluctuations in
Nu around the mean value at early times have both a larger magnitude and higher
frequency than the fluctuations in the relaxed state.
In Fig.~\ref{fig:nu_v_time}b, we show the AE traces of Nu and
$\angles{T}-T_{\text{top}}$ at $S = 10^5$. The times at which AE solves occur are marked
by arrows. Nu reaches its evolved mean value slightly more rapidly than in the SE case,
and the frequency and magnitude of fluctuations in Nu away from the mean resemble the
final relaxed state of SE.

@@

$10^{1/3}$	    &	$2.79 \times 10^3$       &	32$\times$64$\,\,\,\,\,\,\,$&	$52.8$	&	$340$	&	100	&	1.46	&	1.46	\\
$10^{2/3}$	    &	$6.01 \times 10^3$       &	32$\times$64$\,\,\,\,\,\,\,$&	$77.6$	&	$282$	&	100	&	1.95	&	1.95	\\
$10^1$	        &	$1.30 \times 10^4$       &  	32$\times$64$\,\,\,\,\,\,\,$&	$114$	&	$265$	&	100	&	2.43	&	2.42	\\
